Modal Abstractions for Virtualizing Memory Addresses

### 2023-04-05 programlogic.tex
+

Given the definition of physical page-pointsto assertion and the root address of virtual-address space kept in the control register \textsf{CR3} as shown in Figure \ref{fig:pagetable},  one can build the physical address-translation for a virtual address (e.g. \textsf{va}) via abstracting the L4-L1 table traversal as the following:

@@

  \item Page Map Level-4 Translation (L4): Performs the first level address translation to get the PML4 entry (PML4E) in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\textsf{CR3}) to get the starting address of the PML4 table (PLM4T) and using the offset mask \textsf{va}$|^{12}$(\textsf{CR3}). It is worth noting that our offset indices are 12 bits as all physical memory pointsto assertions has to satisfy being \textsf{\textit{aligned}}, e.g. [47-39] plus 3 alignment bits in Figure \ref{fig:pagetables}

@@

@@

\begin{figure}
\begin{mathpar}
\inferrule[WriteToRegFromReg]{
  \{P \ast r_d \mapsto_{r} \textsf{rvs} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}\;\overline{ is}
}{
  \{P \ast r_d \mapsto_{r} \textsf{rvd} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}
  \textsf{ mov}~\textsf{r}_d,~\textsf{r}_s;\;\overline{is}
  %\lstinline|mov %cr3, r| 
}
\\
\inferrule[WriteToPhysMemFromReg]{
  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}\;\overline{ is}  
}{
  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}   \ast r_a \mapsto_{r}\{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{rvs} \}
  \textsf{ mov}~\textsf{r}_a,~\textsf{r}_s;\;\overline{is}
}
\\
\inferrule[WriteToRegFromPhysMem]{
  \{P \ast r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}\;\overline{is}
}{
  \{P \ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}
  \textsf{mov}~\textsf{r}_d,~\textsf{r}_a;\;\overline{is}
}
\\
\inferrule[WriteToRegFromVirtMem]{
  \{P \ast r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}\;\overline{is}
}{
  \{P \ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}
  \textsf{mov}~\textsf{r}_d,~\textsf{r}_a;\;\overline{is}
}
\\
\inferrule[WriteToVirtMemFromReg]{
  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}\;\overline{ is}  
}{
  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}   \ast r_a \mapsto_{r}\{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{rvs} \}
  \textsf{ mov}~\textsf{r}_a,~\textsf{r}_s;\;\overline{is}
}
\\
\inferrule[WriteToRegCtlFromReg]{
  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}  \}\overline{is}
}{
  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}   \}
  \textsf{mov}~\textsf{cr3},~r_s;\;\overline{is}
  %\lstinline|mov %cr3, r| 
}\\
\inferrule[WriteToRegFromRegCtl]{
  \{P \ast r_d \mapsto_{r} \textsf{rvs} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}\;\overline{ is}
}{
  \{P \ast r_d \mapsto_{r} \textsf{rvd} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}
  \textsf{ mov}~\textsf{r}_d,~\textsf{r}_s;\;\overline{is}
  %\lstinline|mov %cr3, r| 
}
\end{mathpar}
\caption{Reasoning Rules for Selected \textsf{AMD64} Instructions}
\label{fig:wpdamd}
\end{figure}
\subsubsection{Register To Memory}
\paragraph{Register To Physical Memory}
\paragraph{Register To Virtual Memory}
\subsubsection{Memory To Register}
\paragraph{Physical Memory To Register}
\paragraph{Virtual Memory To Register}
\subsubsection{Register To Register}
\paragraph{Updating \lstinline|cr3|}

>>>>

Given the definition of physical page-pointsto assertion and the root address of virtual-address space kept in the control register \lstinline{cr3} as shown in Figure \ref{fig:pagetable},  one can build the physical address-translation for a virtual address (e.g. \textsf{va}) via abstracting the L4-L1 table traversal as the following:

@@

  \item Page Map Level-4 Translation (L4): Performs the first level address translation to get the PML4 entry (PML4E) in Figure \ref{fig:pagetables} by using the masks(\textsf{va}$|^{52}$(\lstinline{cr3}) to get the starting address of the PML4 table (PLM4T) and using the offset mask \textsf{va}$|^{12}$(\lstinline{cr3}). It is worth noting that our offset indices are 12 bits as all physical memory pointsto assertions has to satisfy being \textsf{\textit{aligned}}, e.g. [47-39] plus 3 alignment bits in Figure \ref{fig:pagetables}

@@

  \begin{figure*}
\[
\begin{array}{l}
  \mathcal{I}\textsf{ASpace} \stackrel{\triangle}{=} \lambda \; \crthree \ldotp
  \exists\;\gammaPred \;\ldotp\; \ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}} \ast \\
  \bigast{(\vaddr, \entryo)\in \ptablestore}{\exists\;(\entryf\;,\entrytr\;,\entrytw)\ldotp \textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\crthree,\entryf,\entrytr,\entrytw,\entryo)}
\end{array}
\]
\caption{Global Address-Space Invariant}
  \label{fig:peraspaceinvariant}
  \end{figure*}

@@

\begin{figure}
\begin{mathpar}
\inferrule[WriteToRegFromReg]{
  \{P \ast r_d \mapsto_{r} \textsf{rvs} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}\;\overline{ is}
}{
  \{P \ast r_d \mapsto_{r} \textsf{rvd} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}
  \textsf{ mov}~\textsf{r}_d,~\textsf{r}_s;\;\overline{is}
  %\lstinline|mov %cr3, r| 
\\
\inferrule[WriteToPhysMemFromReg]{
  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}\;\overline{ is}  
}{
  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}   \ast r_a \mapsto_{r}\{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{rvs} \}
  \textsf{ mov}~\textsf{r}_a,~\textsf{r}_s;\;\overline{is}
\\
\inferrule[WriteToRegFromPhysMem]{
  \{P \ast r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}\;\overline{is}
}{
  \{P \ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v} \}
  \textsf{mov}~\textsf{r}_d,~\textsf{r}_a;\;\overline{is}
\\
\inferrule[WriteToRegFromVirtMem]{
  \{P \ast r_d \mapsto_{r}  \textsf{v} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}\;\overline{is}
}{
  \{P \ast r_d \mapsto_{r}  \textsf{rvd} \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}
  \textsf{mov}~\textsf{r}_d,~\textsf{r}_a;\;\overline{is}
\\
\inferrule[WriteToVirtMemFromReg]{
  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}  \ast r_a \mapsto_{r} \{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{v} \}\;\overline{ is}  
}{
  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{rvs}   \ast r_a \mapsto_{r}\{q\} \textsf{ vaddr} \ast \textsf{vaddr} \mapsto_{\textsf{v,rtv}} \textsf{rvs} \}
  \textsf{ mov}~\textsf{r}_a,~\textsf{r}_s;\;\overline{is}
\\
\inferrule[WriteToRegCtlFromReg]{
  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}  \}\overline{is}
}{
  \{P \ast r_s \mapsto_{r}\{q\}  \textsf{ rvs}   \}
  \textsf{mov}~\textsf{cr3},~r_s;\;\overline{is}
  %\lstinline|mov %cr3, r| 
}\\
\inferrule[WriteToRegFromRegCtl]{
  \{P \ast r_d \mapsto_{r} \textsf{rvs} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}\;\overline{ is}
}{
  \{P \ast r_d \mapsto_{r} \textsf{rvd} \ast r_s \mapsto_{r}\{q\} \textsf{ rvs} \}
  \textsf{ mov}~\textsf{r}_d,~\textsf{r}_s;\;\overline{is}
  %\lstinline|mov %cr3, r| 
\end{mathpar}
\caption{Reasoning Rules for Selected \textsf{AMD64} Instructions}
\label{fig:wpdamd}
\subsubsection{Memory To Register}
In Figure \ref{fig:wpdamd}, we see the rules for updating the memory address ($r_a$). To do so, we essentially need page-table-walk pointsto ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) to locate the value to be written. We mainly have two options to achieve this. 
\paragraph{Physical Memory To Register} First one is to use  page-table-walk pointsto ($\textsf{vaddr} \mapsto_{\textsf{t}} \textsf{v}$), which is consists of all-physical page-table memory assertions, to prove the page-table traversal \textsf{AMD64} implements, update the register $r_d$ the value obtained ($\textsf{v}$) as shown in Figure (\TirNameStyle{WriteToRegFromPhysMem}). By doing so, we expose the resources ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$) needed only for the proof of \textsf{AMD64} page-table traversal to the client of the rule. This breaks the modularity of reasoning we want to maintain in location virtualization.
  \paragraph{Virtual Memory To Register}  Therefore, we would like to use a virtual-pointsto assertion ($\textsf{vaddr} \mapsto_{\textsf{v},\textsf{rtv}}$), which hides physical resources needed for the proof of page-table-traversal from the client, to obtain the the value ($\textsf{v}$) to update the register $r_d$ as shown in Rule (\TirNameStyle{WriteToRegFromVirtMem}). While doing the proof of page-table-traversal with a given virtual-pointsto definition in Figure \ref{fig:virtualpointstosharing}, we unfold and obtain our token ($\sumwalkabs\vaddr\qfrac\paddr$) to exchange for a physical-table mapsto assertion inside the $\mathcal{I}\textsf{ASpace}$ ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$) in Figure \ref{fig:peraspaceinvariant}, and apply Rule (\TirNameStyle{WriteToRegFromPhysMem}).
\subsubsection{Register To Memory}
 \paragraph{Register To Physical Memory} First one is to

\paragraph{Register To Virtual Memory}
\subsubsection{Register To Register}
\paragraph{Updating \lstinline|cr3|}
